Regression

### 2015-08-27 rand_forest.rmd
tried boosted trees

heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)

@@

addone<- function(var= 'JD.Second', df= all7d){
        reduced = df[!(is.na(df[,var])),]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng", var, sep=" + ")
        add.one = rpart(as.formula(formula), data = reduced)

@@

@@

addone(var='JD.Second', df=all7d)
#RMSE = .528
addone(var='Post.Code', df=all7d)
#RMSE = .499
addone(var='majority.pos', df=all7d)
#RMSE = .49
addone(var='pc.majpos', df=all7d)
#RMSE = .486
addone(var='code.client', df=all7d)
#RMSE = .534
addone(var='Billing.Type', df=all7d)
#RMSE = .502
addone(var='code.contact', df=all7d)
#RMSE = .535

@@

addmult<- function(var= c('JD.Second', 'code.client'), combos= c('JD.Second*inv.mlsto.log'), df= all7d){
        reduced = df[complete.cases(df[,var]),]
        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business")

@@

        for(i in 1:length(combos)){
                formula = paste(formula, combos[i], sep= " + ")
        }

        add.one = lm(as.formula(formula), data = reduced)

        print(summary(add.one))

        ans= data.frame(answer= reduced$return.pdol, addone.lm = predict(add.one, reduced, interval= 'predict'))

        print(head(ans))

        print(
        paste("RMSE=", RMSE(ans$answer, ans$addone.lm.fit) %>% round(4), sep= " ")
        )

@@

#try Billing.Type with inv.mlsto.log
addmult(var= c('Billing.Type'), combos= c('Billing.Type*inv.mlsto.log'), df= all7d)
# no

#try code.client*inv.mlsto.log
addmult(var= c('code.client'), combos= c('code.client*inv.mlsto.log'), df= all7d)
#made a difference, brought RMSE down to .4076 :)!

#try JD.Second with inv.mlsto.log (no), code.client (too intense), timespan (no), pc.pro (no), no.users (yes)?
addmult(var= c('JD.Second'), combos= c('JD.Second*no.users'), df= all7d)
#add JD.Second*no.users

addmult(var= c('majority.pos'), combos= c('majority.pos*inv.mlsto.log'), df= all7d)
#no

addmult(var= c('pc.majpos.log'), combos= c('pc.majpos.log*inv.mlsto.log'), df= all7d)
#no

@@

It is worthwhile to include:

* code.client by inv.mlsto.log because brought down RMSE from 0.4394 (code.client) to 0.4076 (code.client + code.client by inv.mlsto.log)
* JD.Second by no.users. Brought down RMSE from 0.4709 to 0.4612...

Now, how to get variable importance from lm..

@@

#variable importance for full variables only - pairs.lm
#use varImp from caret package - uses t-test statistic

imp<- varImp(pairs.lm, scale=FALSE)
imp$variable<- rownames(imp)
imp<- imp %>% select(variable, Overall)
imp %>% arrange(-Overall) %>% head()

sum(imp$Overall)

>>>>

heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)

@@

addone<- function(var= 'JD.Second', min.Year = 2008, df= all7d){
        reduced = df[!(is.na(df[,var])) & df$Year > min.Year,]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type", var, sep=" + ")
        add.one = rpart(as.formula(formula), data = reduced)

@@

        print(
        train.rpart$results
        )
}

@@

addone(var='JD.Second', min.Year = 2008, df=all7d)
#RMSE = .528
addone(var='Post.Code', min.Year = 2008, df=all7d)
#RMSE = .499
addone(var='majority.pos', min.Year = 2008, df=all7d)
#RMSE = .49
addone(var='pc.majpos', min.Year = 2008, df=all7d)
#RMSE = .486
addone(var='code.client', min.Year = 2008, df=all7d)
#RMSE = .534
addone(var='Billing.Type', min.Year = 2008, df=all7d)
#RMSE = .502
addone(var='code.contact', min.Year = 2008, df=all7d)
#RMSE = .535

@@

addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type")

@@

        add.mult = rpart(as.formula(formula), data = reduced)

        #select cp
        cp.new = add.mult$cptable[which.min(add.mult$cptable[,"xerror"]),"CP"]-0.0002
        add.mult = rpart(as.formula(formula), data = reduced,
                        control = rpart.control(cp= cp.new))

        #plot decision tree
        print(
          heat.tree(add.mult, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
          )

@@

#try Billing.Type with inv.mlsto.log
## YES! print this one!! 
addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
#RMSE at 48.6 cents

@@

Adding aditional variables is making everything worse.
Try randomForest

```{r, echo=FALSE}
#perform 10 fold xval using caret package
tc = trainControl("cv",5)
rf.grid = expand.grid(mtry= 1:10)
min.form= "return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type"
train.rf = train(as.formula(min.form), data=all7d, method="rf",trControl=tc,
                    tuneGrid=rf.grid)
#result - use mtry= 5

#make first single randomForest
first.rf<- randomForest(as.formula(min.form),
                  data=all7d,mtry=1,ntree=500,importance=TRUE)

#already have made train and test set
#try to predict medv
rf.all<-randomForest(as.formula(min.form),data=all7d,mtry=1,ntree=500,importance=TRUE)

#create vectors full of '0's
oob.err<-double(7)
test.err<-double(7)
for(mtry in 1:7){
        fit<-randomForest(as.formula(min.form),data=train, mtry=ntry, ntree=500)
        oob.err[mtry]<-fit$mse[mtry]
        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2) %>% sqrt()
        cat(mtry," ")
}

qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
        geom_line(aes(1:mtry,oob.err),colour="blue")+
        geom_point(aes(1:mtry,oob.err),colour="blue")

#mtry = 3 is optimal! test RMSE of 48.2 cents ....
oob.err<-double(5)
test.err<-double(5)
ntree = c(300,400,500,600,700)
for(i in 1:5){
        fit<-randomForest(as.formula(min.form),data=train, mtry=3, ntree=ntree[i])
        oob.err[i]<-fit$mse[i]
        test.err[i]<- mean((predict(fit,test)-test$return.pdol)^2) %>% sqrt()
        cat(i," ")
}

qplot(ntree,test.err,geom=c("point","line"),color="pink")+
        geom_line(aes(ntree,oob.err),colour="blue")+
        geom_point(aes(ntree,oob.err),colour="blue")
#doesn't really matter, use 400!

#final forest with full variables

full.forest<- randomForest(as.formula(min.form),data=all7d,mtry=5,ntree=500,importance=TRUE)
plot(full.forest)
importance(full.forest)
varImpPlot(full.forest)

full.forest<- randomForest(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type +code.client, data= all7d[complete.cases(all7d$code.client),],
                  mtry=5,ntree=500,importance=TRUE)

Try adding variables

@@

addmult<- function(var= c('JD.Second', 'code.client'), min.Year = 2008, df= all7d){
        reduced = df[complete.cases(df[,var]) & df$Year>min.Year,]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type")

        add.mult = randomForest(as.formula(formula),mtry=5, ntree=500, data = reduced)

        #perform 10 fold xval using caret package
        tc = trainControl("cv",5)
        #keep cp static at 0.011
        rf.grid = expand.grid(mtry= 5:6)
        train.rf = train(as.formula(formula), data=df, method="rf",trControl=tc,
                  tuneGrid=rf.grid)

        print(
        train.rf
        )

#try Billing.Type with inv.mlsto.log
## YES! print this one!! 
addmult(var= c('pc.majpos', 'majority.pos'), min.Year = 2008, df= all7d)
#RMSE at 48.6 cents
